---
title: "Assignment 4"
author: "Laura Fusi"
date: "2023-03-17"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(rtracklayer)
  library(epiwraps)
  library(GenomicRanges)
})
ah <- AnnotationHub()
```

# Find and download from ENCODE the peaks (i.e. bed-like format) for the following histone modifications and transcription factors in mouse embryonic stem cells (mESC):
ENCODE: https://www.encodeproject.org/search/?type=Experiment&control_type!=*&status=released 
(when there are replicates, we recommend using the bed file denoted as “conservative IDR thresholded peaks”)

## p300
https://www.encodeproject.org/files/ENCFF653YVR/

```{r}
download.file("https://www.encodeproject.org/files/ENCFF653YVR/@@download/ENCFF653YVR.bed.gz", dest="p300.bed.gz")
p300_peaks <- rtracklayer::import("p300.bed.gz", format="narrowPeak")
p300_peaks
```
There are in total 28341 peaks.

## H3K4me3

https://www.encodeproject.org/files/ENCFF247GVM/

```{r}
download.file("https://www.encodeproject.org/files/ENCFF247GVM/@@download/ENCFF247GVM.bed.gz", dest="H3K4me3.bed.gz")
H3K4me3_peaks <- rtracklayer::import("H3K4me3.bed.gz", format="narrowPeak")
```

## H3K4me1

https://www.encodeproject.org/files/ENCFF333IJH/

```{r}
download.file("https://www.encodeproject.org/files/ENCFF333IJH/@@download/ENCFF333IJH.bed.gz", dest="H3K4me1.bed.gz")
H3K4me1_peaks <- rtracklayer::import("H3K4me1.bed.gz", format="narrowPeak")
```

## H3K27ac

https://www.encodeproject.org/files/ENCFF360VIS/

```{r}
download.file("https://www.encodeproject.org/files/ENCFF360VIS/@@download/ENCFF360VIS.bed.gz", dest="H3K27ac.bed.gz")
H3K27ac_peaks <- rtracklayer::import("H3K27ac.bed.gz", format="narrowPeak")
```

## H3K27me3

https://www.encodeproject.org/files/ENCFF008XKX/ 

```{r}
download.file("https://www.encodeproject.org/files/ENCFF008XKX/@@download/ENCFF008XKX.bed.gz", dest="H3K27me3.bed.gz")
H3K27me3_peaks <- rtracklayer::import("H3K27me3.bed.gz", format="narrowPeak")
```

# Of the p300 peaks, what proportion overlap each of the marks?

## H3K4me3

```{r}
p300OverlapsH3K4me3 <- overlapsAny(p300_peaks, H3K4me3_peaks)
table(p300OverlapsH3K4me3)
```
Of the 24596 p300 peaks, 8742 overlap with H3K4me3 marks.

```{r}
table(p300OverlapsH3K4me3)[2]/(table(p300OverlapsH3K4me3)[1]+table(p300OverlapsH3K4me3)[2])
```
This corresponds to ca. 35.5%.

## H3K4me1

```{r}
p300OverlapsH3K4me1 <- overlapsAny(p300_peaks, H3K4me1_peaks)
table(p300OverlapsH3K4me1)
```
Of the 24596 p300 peaks, 6905 overlap with H3K4me1 marks.

```{r}
table(p300OverlapsH3K4me1)[2]/(table(p300OverlapsH3K4me1)[1]+table(p300OverlapsH3K4me1)[2])
```
This corresponds to ca. 28.1%.

## H3K27ac

```{r}
p300OverlapsH3K27ac <- overlapsAny(p300_peaks, H3K27ac_peaks)
table(p300OverlapsH3K27ac)
```
Of the 24596 p300 peaks, 13502 overlap with H3K27ac marks.

```{r}
table(p300OverlapsH3K27ac)[2]/(table(p300OverlapsH3K27ac)[1]+table(p300OverlapsH3K27ac)[2])
```
This corresponds to ca. 54.9%.

## H3K27me3

```{r}
p300OverlapsH3K27me3 <- overlapsAny(p300_peaks, H3K27me3_peaks)
table(p300OverlapsH3K27me3)
```
Of the 24596 p300 peaks, 56 overlap with H3K27me3 marks.

```{r}
table(p300OverlapsH3K27me3)[2]/(table(p300OverlapsH3K27me3)[1]+table(p300OverlapsH3K27me3)[2])
```
This corresponds to ca. 0.2%.

# Session information

```{r}
sessionInfo()
```

